---
title: "ETC3250 kaggle comp"
author: "Cameron Roach"
date: "13/05/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())

library(tidyverse)
library(caret)

load("data/ao_training.rda")
load("data/ao_test_unlabelled.rda")

```

# Submission

## Process the data

```{r process-data}
process_data <- function(x, data_name = "data") {
  x <- x %>% 
    as_tibble() %>% 
    mutate(
      gender = case_when(
        str_sub(matchid, 1, 2) == "MS" ~ "M",
        str_sub(matchid, 1, 2) == "WS" ~ "F",
        TRUE ~ str_sub(matchid, 1, 2))
    )
  
  df_1 <- x %>% 
    group_by(pointid) %>% 
    filter(n() == 1) %>% 
    ungroup() %>% 
    select(-c(event, year, matchid)) %>% 
    mutate_if(is.factor, as.character) %>% 
    select(pointid, matches("outcome"), gender, everything()) %>% 
    # The following variables only have one value (because 1 serve rally)
    # so they are removed. Prevents error in caret::train.
    select(-c(starts_with("previous"), shotinrally, rally, 
              server.is.impact.player, outside.baseline, hitpoint)) %>% 
    mutate(net.clearance = if_else(net.clearance == -0.5, 1.5, net.clearance))
  
  df_2 <- x %>% 
    group_by(pointid) %>% 
    filter(n() == 2) %>% 
    select(-c(event, year, matchid)) %>% 
    arrange(pointid, shotinrally) %>% 
    mutate(idx = 1:n()) %>% 
    ungroup() %>% 
    mutate_all(as.character) %>%  # prevents warning in gather
    gather(var, val, -c(pointid, gender, matches("outcome"), idx, serve, serverwon)) %>%
    unite(var_idx, var, idx) %>% 
    spread(var_idx, val) %>% 
    select(-previous.speed_2) %>%  # same as speed_1
    rename(speed_0 = previous.speed_1) %>%  # rename to be consistent with speed_1, speed_2
    mutate_at(vars(
      starts_with("serve_"),
      starts_with("speed_"),
      starts_with("net.clearance_"),
      starts_with("time.to.net_"),
      starts_with("distance.from.sideline_"),
      starts_with("depth_"),
      starts_with("player.impact.distance.travelled_"),
      starts_with("player.impact.depth_"),
      starts_with("player.impact.distance.from.center_"),
      starts_with("previous.speed_"),
      starts_with("previous.net.clearance_"),
      starts_with("previous.distance.from.sideline_"),
      starts_with("previous.depth_"),
      starts_with("opponent.distance.travelled_"),
      starts_with("opponent.depth_"),
      starts_with("opponent.distance.from.center_"),
      starts_with("previous.time.to.net_"),
      starts_with("shotinrally_"),
      starts_with("rally_")
    ), as.numeric) %>% 
    mutate_at(vars(
      starts_with("outside.sideline_"),
      starts_with("outside.baseline_"),
      starts_with("same.side_"),
      starts_with("server.is.impact.player_")
    ), as.logical) %>% 
    mutate_if(is.factor, as.character) %>% 
    mutate(ball_dist_x_11 = opponent.distance.from.center_1 + distance.from.sideline_1,
           ball_dist_x_12 = opponent.distance.from.center_1 + distance.from.sideline_2,
           ball_dist_x_21 = opponent.distance.from.center_2 + distance.from.sideline_1,
           ball_dist_x_22 = opponent.distance.from.center_2 + distance.from.sideline_2,
           ball_dist_y_11 = opponent.depth_1 - depth_1,
           ball_dist_y_12 = opponent.depth_1 - depth_2,
           ball_dist_y_21 = opponent.depth_2 - depth_1,
           ball_dist_y_22 = opponent.depth_2 - depth_2,
           net_clearance_12 = net.clearance_1 + net.clearance_2,
           opponent_distance_opp_impact = player.impact.distance.from.center_2 + opponent.distance.from.center_1,
           speed_change_21 = speed_2 - speed_1,
           speed_change_10 = speed_1 - speed_0,
           speed_total_21 = speed_2 + speed_1,
           speed_total_10 = speed_1 + speed_0,
           speed_total_210 = speed_2 + speed_1 + speed_0) %>% 
    select(pointid, matches("outcome"), gender, everything())
  
  
  
  tibble(n_shots = c(1, 2),
         !!data_name := list(df_1, df_2))
}

model_df <- process_data(ao_tr, "train_data") %>% 
  inner_join(process_data(ao_ts_unlabelled, "test_data"), by = "n_shots")
```


```{r fit-models}
xgb_ctrl <- trainControl(
  method = "cv",
  number = 10,
  allowParallel = FALSE,
  returnData = FALSE,
  trim = TRUE,
  returnResamp = "none",
  savePredictions = "none"
)


fit_xgb_linear_model <- function(data, shot) {
  cat("Fitting xgbLinear model for", shot, "shot rallies...\n")
  
  data %>%
    select(-pointid) %>%
    na.omit() %>%
    train(outcome ~ .,
          data = .,
          method="xgbLinear",
          metric="Accuracy",
          trControl = xgb_ctrl,
          nthread = 11,
          tuneGrid = list(nrounds = 100,
                          lambda = c(0, exp(-c(10, 8, 6, 4, 2, 0))),
                          alpha = 0,
                          eta = 0.2) %>%
            cross_df() %>%
            as.data.frame())
}


fit_xgb_tree_model <- function(data, shot) {
  cat("Fitting xgbTree model for", shot, "shot rallies...\n")

  data %>%
    select(-pointid) %>%
    na.omit() %>%
    train(outcome ~ .,
          data = .,
          method="xgbTree",
          trControl = xgb_ctrl,
          tuneGrid = list(nrounds = 150,
                          max_depth = 3:6,
                          eta = c(0.2, 0.3),
                          gamma = c(0, 0.1, 0.2, 0.3),
                          colsample_bytree = .8,
                          min_child_weight = 1,
                          subsample = c(.75, 1)) %>%
            cross_df() %>%
            as.data.frame())
}


model_df <- model_df %>% 
  mutate(xgb_linear = map2(train_data, n_shots, fit_xgb_linear_model),
         xgb_tree = map2(train_data, n_shots, fit_xgb_tree_model))

save(model_df, file = "cache/model_df_1.Rdata")
```


## Predict results

```{r}
train_predict <- function(data, object) {
  data %>% 
    mutate(outcome = predict(object, data)) %>% 
    select(pointid, outcome) %>% 
    mutate_all(as.character)
}

prediction_df <- model_df %>% 
  mutate(predictions = map2(test_data, xgb_tree, train_predict)) %>% 
  select(predictions) %>% 
  unnest() %>% 
  arrange(pointid)

write_csv(prediction_df, "cache/predictions_2.csv")
```



# Data exploration

## Feature importance

```{r}
xgb.importance(model = model_df$xgb_tree[[2]]$finalModel) %>% 
  as.data.frame() %>% 
  gather(var, val, -Feature) %>% 
  ggplot(aes(x = Feature, y = val, colour = var)) + 
  geom_col(width= 0) + 
  geom_point() +
  facet_wrap(~var, ncol = 1) + 
  theme(legend.position="none", 
        axis.text.x = element_text(angle = 45, hjust = 1))
```



## Feature engineering ideas

Remember to only look at those classes that are being misclassified. Might already be covering for these ideas with the fitted model.

* net.clearance_2 and speed_2 could be a good predictor of unforced errors. Too fast and too high => goes out.
* net.clearance_2 and distance
* Check this for some ideas https://github.com/ml874/Predicting-How-Points-End-in-Tennis/blob/master/Gradient%20Boosting.ipynb
* change in player distance (if somebody rushes the net they might make an error). Change between 2-1 and 1-0.

```{r}
# model_df$xgb_linear[[2]]$results %>% 
#   select(nrounds:Accuracy) %>% 
#   gather(var, val, -Accuracy) %>% 
#   filter(var == "lambda") %>% 
#   ggplot(aes(x = val, y = Accuracy)) + 
#   geom_point() +
#   scale_x_log10()# + 
#   #geom_smooth()

model_df$xgb_tree[[2]]$results %>% 
  select(eta:Accuracy) %>% 
  gather(var, val, -Accuracy) %>% 
  ggplot(aes(x = val, y = Accuracy, group = val)) + 
  geom_boxplot() + 
  facet_wrap(~var, scales = "free_x")

3942/(3942+22986)*.99997+22986/(3942+22986)*0.9167317
```



```{r}
pred_class_df <- model_df$data[[2]] %>% 
  mutate(pred = predict(model_df$xgb_tree[[2]], model_df$data[[2]]))

table(pred_class_df$outcome, pred_class_df$pred)

table(pred_class_df$outcome, pred_class_df$pred) %>% 
  as.data.frame() %>% 
  group_by(Var1) %>% 
  mutate(accuracy = Freq/sum(Freq)) %>% 
  filter(Var1 == Var2)
```


```{r}
bad_df <- pred_class_df %>% 
  filter(outcome %in% c("F", "U"),
         outcome != pred)
  
# bad_df %>% 
  # mutate() %>%
  # select(gender, outcome, starts_with("ball_dist"), net_clearance_12, opponent_distance_opp_impact, speed_change_21) %>%
  # select(gender, outcome, speed_change_21, speed_total_21) %>% 
  # gather(var, val, -c(outcome, matches("gender"))) %>% 
  # ggplot(aes(x = val, colour = outcome)) +
  # geom_jitter()
  # geom_density() +
  # geom_rug() +
  # facet_grid(gender~var)
  # facet_wrap(~var)


bad_df %>% 
  filter(outcome %in% c("F", "U")) %>% 
  ggplot(aes(x = player.impact.distance.travelled_2, y = speed_2, colour = outcome)) +
  geom_point() +
  facet_wrap(~gender)
```




```{r}
library(GGally)

pairs_df <- bad_df %>% 
  select_if(~ is.numeric(.x) | is.logical(.x)) %>% 
  mutate(outcome = bad_df$outcome) %>% 
  select(outcome, everything())


#pairs_df[sample(1:dim(pairs_df)[1], 200), sample(1:dim(pairs_df)[2], 10)] %>% 
pairs_df %>% 
  sample_frac(0.1) %>% 
  select_at(c(1, 2:10)) %>% 
  ggpairs(aes(colour = outcome), shape = "o", alpha = 0.5)
```




